---
title: "Data Visualization"
author: "Ali Paulson"
date: "February 14, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
library(leaflet)
library(DT)
```

The data we are using is from [Alaska Department of Fish and Game, Division of Commercial Fisheries. 2017. Daily salmon escapement counts from the OceanAK database, Alaska, 1921-2017. Knowledge Network for Biocomplexity. doi:10.5063/F1S46Q6M.]("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e")


#Reformat the data for plotting
```{r}
salmon <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e", method = "libcurl"), stringsAsFactors = F)

head(salmon)
```

* Calculate annual escapment by species and regoin
    + `separate`
    + `group_by`
    + `summarize` using `sum` function

* filter for just Pacific salmon species (Sockeye, Chinook, Coho, Pink, Chum)

```{r}
annual_esc <- salmon %>% 
  separate(sampleDate, into = c("year", "month", "day"), sep = "-") %>% 
  group_by(year, SASAP.Region, Species) %>% 
  summarize(escapement = sum(DailyCount)) %>% 
  filter(Species %in% c("Sockeye", "Chinook", "Coho", "Pink", "Chum")) %>% 
  ungroup() %>% 
  #coerce year to numeric type: 
  mutate(year = as.numeric(year))

head(annual_esc)
```


#Generate some static plots

```{r}
#Better practice to put your mapping in the ggplot call rather than the geom call, unless you are trying to map different parameters on the same plot

ggplot(annual_esc, mapping = aes(x = Species, y = escapement))+
  geom_col(fill = "blue")

#Check out other color palettes with RColorBrewer

#What if you do want to map something with the fill aesthetic?
ggplot(annual_esc, mapping = aes(x = Species, y = escapement, fill = SASAP.Region))+
  scale_fill_brewer(palette = 1)+ #Can choose different palettes for the fill
  geom_col()

```


What if we just wanted to plot the Kodiak escapement?

```{r}
annual_esc %>% 
  filter(SASAP.Region == "Kodiak") %>% 
  ggplot(aes(x = year, y = escapement, color = Species))+
  geom_line()+
  geom_point()+
  ggtitle("Kodiak Salmon Escapement")+
  theme_test() + #Can use different ggplot themes to make things prettier
  theme(legend.position = "bottom", legend.title = element_blank())

#These theme calls for ggplot can get really long, but you can save your theme function calls to an object: 

my_theme = theme_test()+
  theme(legend.position = "Bottom", legend.title = element_blank())

#This makes your code more dry, less repetition, easier to read. 
#You want to avoid calling theme twice though - as you can easily overwrite some of the things that you set as your theme - so if you have a theme() in your my_theme, you might overwrite things - just be careful. 
#Use scale package to fix up the axis labels

annual_esc %>% 
  filter(SASAP.Region == "Kodiak") %>% 
  ggplot(aes(x = year, y = escapement, color = Species))+
  geom_line()+
  geom_point()+
  scale_y_continuous(labels = comma)+
  ggtitle("Kodiak Salmon Escapement")+
  xlab("Year") +
  ylab("Escapement (num of fish)") +
  my_theme #Add in the mytheme object!

```

Let's make multiple plots, one for each region. 

```{r, fig.width=7, fig.height = 8}

p <- ggplot(annual_esc, aes(x = year, y = escapement, color = Species))+
  geom_line()+
  geom_point()+
  scale_y_continuous(labels = comma)+
  facet_wrap(~SASAP.Region, scales = "free_y", ncol =2)+
  ggtitle("Kodiak Salmon Escapement")+
  xlab("Year") +
  ylab("Escapement (num of fish)") +
  my_theme #Add in the mytheme object!

p
ggsave("figures/regional_escapement.png", 
       plot = p, width = 7, height = 8, units = "in")
```


Can also turn some parts of R chunks on or off



#Generate an interactive map using leaflet

Map the sampling locations for escapement. Starting back with the unsummarized data. 

* Find distinct lat/long/location combinations, using the `distinct`
* the `drop_na` function removes whole rows that contain an NA anywhere - could be dangerous, so be careful it's doing what you think. 

```{r, echo = FALSE}
locations <- salmon %>% 
  distinct(Location, Latitude, Longitude) %>%  #pass this function the variables that you want it to pull distinct values from
  drop_na()

head(locations)

datatable(locations)
```

Let's make a map!

* use pipes with the leaflet package, rather than plus signs
* Leaflet has a ton of functionality - can create an html block within the popups, can put in plots/hyperlinks into the popups
* This isn't greaet for developing static maps - a new package called sf  - this integrates with ggplot2 for mapping vector files.  
* Can create custom markers in leaflet as well. 

```{r}
leaflet(locations) %>% 
  addTiles() %>% #Can change projection, but will have to use different tiles - default tiles use mercator
  addMarkers(lng = ~Longitude, lat = ~Latitude, popup = ~Location)
```


If you want to add breaks to your html output: 
adfljsa

<br/>
asdflkj