---
title: "functions-tutorial"
author: "Ali Paulson"
date: "February 13, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Work with some data in degrees F

- Convert to C

By doing it this way, you can easily introduce errors
```{r}
airtemps <- c(212, 100, 32, 64, 0, -20)
airtemps
airtemp1 <- (airtemps - 32) *5/9
airtemp1
```


# Write a function to convert to C

You use the function() function to create a function. 

* temp_fahr is the input
* the code in between the braces will be run, and will be assigned the value fahr_to_celsius
* you use return to return a value back...R's default is to return the last value in your function, so you should be explicit about what you want returned

```{r}
fahr_to_celsius <- function(temp_fahr) {
  celsius <- (temp_fahr - 32) * 5/9
  return(celsius)
}


fahr_to_celsius(32)
fahr_to_celsius(78)
celsius_temps <- fahr_to_celsius(airtemps) #Works like a vector function, too
celsius_temps
fahr_to_celsius

```
Can type function name (without parentheses) into the console to see the code behind a function. 

Can provide defaults in your argument calls...


```{r}
fahr_to_celsius_default <- function(temp_fahr = 100) {
  celsius <- (temp_fahr - 32) * 5/9
  return(celsius)
}


fahr_to_celsius_default()
fahr_to_celsius_default(32)

```


# Write a function to convert C to F


```{r}
celsius_to_fahr <- function(temp_C){
  fahr <- (temp_C * 9/5)+32
  return(fahr)
}

celsius_to_fahr(0)

fahr_temps <- celsius_to_fahr(celsius_temps)
fahr_temps
setdiff(fahr_temps, airtemps)
setdiff(airtemps, fahr_temps)
fahr_temps == airtemps

```

# Show how to document functions

You should document functions after you write them 

* use a package called roxygen
```{r}
#' Converts numeric temperature data from degrees Fahrenheit to degrees Celsius
#' 
#' @param temp_fahr is the value in F to be converted to C
#' @return celsius is the temperature value in C, converted from F

fahr_to_celsius <- function(temp_fahr) {
  celsius <- (temp_fahr - 32) * 5/9
  return(celsius)
}


```

#Can create an R Code (not markdown) that contains all of the functions that you might want to use in different codes.  Then, you can **source()** that file be able to use those functions in your code. You can also build your own package that holds all of your functions. 

I've created an R code where I included the two functions (annotated with a new to distinguish from above). Then, I will source this. 

```{r}
source("../Code/conversions.R")
airtemps
a <- new_fahr_to_celsius(airtemps)
a
new_celsius_to_fahr(a)
```

Now, when you have a whole bunch of functions, you have to decide where to store them in your file system.  Tomorrow, we will talk about putting these functions together into a package, so that you can find them wherever you are. 


