---
title: "TidyData"
author: "Ali Paulson"
date: "February 13, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
#Can use warning = FALSE, message = FALSE to suppress warnings in the html output
```


The data we are using are from [Mike Byerly. Alaska commercial salmon catches by management region (1886- 1997). Gulf of Alaska Data Portal. df35b.304.2.](https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1) This is from a PDF file, retrieved with OCR - optical character recognition. You can use this to get a computer to read a PDF. 

```{r data, echo = FALSE}
#catch <- read.csv("../salmondata/df35b.302.1.csv", stringsAsFactors = FALSE)

#regdefs <- read.csv("../salmondata/df35b.303.1.csv", stringsAsFactors = FALSE)

#Can also just download these from the internet!
catch_original <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method = "libcurl"), stringsAsFactors = FALSE)
```

# Reformat Catch Data

```{r}
head(catch_original)
```

Remove the `all` and `notesRegCode` columns

Some people just remove columns by setting them to NULL, but this isn't very clean/readable.  The dplyr world makes things more readable, etc. So, we can use **select()** to remove columns. 

The pipe operator takes the output from one function and puts it into the input of the next function. <- is kind of like "gets the value", %>% is like "gets passed on to/then". Piping makes your workflow more legible, easier for others to understand. 

Ctrl + shift + m: pipe operature shortcut %>% 

Need to explain why you don't have to write in the dataframe as the first argument as select...pipe operator passes things into that first argument. 

**gather()** collects things into key-value pairs

* key is the column names in wide format - key specifies what we want to call the new column of old column names
* value is the value in those columns - value is what we want to call those values in the new long dataframe
* gather() will turn all of your columns into these key-value pairs unless you tell it otherwise
    + can tell gather() what not to include (or what to include)
        + e.g. -Region, -Year
        + or e.g. Chinook, Sockeye, Coho, Pink, Chum

```{r}

catch_long <- catch_original %>% 
  select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum)

head(catch_long)

#equivalent to
# catch_longb <- catch_original %>% 
#   select(-All, -notesRegCode)
# head(catch_longb)

#Now let's make it long
catch_long <- catch_original %>% 
  #Select desired columns
  select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum) %>% 
  #Move from wide to long
  gather(key = "species", value = "catch", -Region, -Year)

head(catch_long)
```


The opposite of **gather()** is **spread()**. 

* key - the column names - this is the column that you want to turn into column names  
* value is the column with the values that you want the new df to have

```{r}
catch_wide <- catch_long %>% 
  spread(key = species, value = catch)
head(catch_wide)

```


#Clean up our data

* rename catch to `catch_thousands`
* create a new catch column in units `number_fish`
* catch column needs to be changed to numeric

If you chain together a ton of pipes without explaining them, it can be confusing for others.  So, it's a good idea to break up pipe operators into logical chunks. 

Function **rename()** allows you to easily change column names. 

Function **mutate()** allows you to add a new column to a dataframe, also allows you to change the structure of a column in a dataframe (e.g. change to numeric from character). 

We get an error when we try to change the catch_thousands column into numeric - sometimes this is okay, but you need to investigate the error messages to make sure nothing fishy is going on. 

```{r}
catch_clean <- catch_long %>% 
  rename(catch_thousands = catch) %>% 
  mutate(catch_thousands = as.numeric(catch_thousands))
  #Create a new column - doesn't work because there is something wrong in the catch_thousands column
 # mutate(catch = catch_thousands * 1000)

head(catch_clean)
```
 
 Look for where as.numeric failed
 
There is an "I" (uppercase i) in the dataframe. Go back to the original data source and double check what this value should be.  Here, it should actually be a 1 (one), but the OCR thought it was an I. 

We need to fix this reproducibly, by explaining the error and what we did to fix it. 

```{r}
test <- as.numeric(catch_long$catch)
i <- which(is.na(test))
#Same as 
#i <- which(is.na(test)==TRUE)

#Can see that the 401st row is the problem
catch_long[i, ]


```


Fix the problem!

Use an **if_else** function to turn the I into a 1.  

```{r}
catch_clean <- catch_long %>% 
  rename(catch_thousands = catch) %>% 
  mutate(catch_thousands = ifelse(catch_thousands == "I", 1, catch_thousands)) %>% 
  #Same as this with the %in% operator: 
  #mutate(catch_thousands = ifelse(catch_thousands == %in% c("I"), 1, catch_thousands)) %>% 
  mutate(catch_thousands = as.numeric(catch_thousands)) %>% 
  mutate(catch = catch_thousands * 1000) %>% 
  #Don't want the catch_thousands column any more: 
  select(-catch_thousands)

head(catch_clean)
```


#Split-apply-combine

Calculate summary metrics, e.g. total or mean catch by region. Calculate mean catch by species. 

```{r}
species_mean <- catch_clean %>% 
  group_by(species, Region) %>% 
  summarize(catch_mean <- mean(catch), 
            num_obs <- n())

head(species_mean)
```



* General QA

# Join the Region Definitions 

# Miscellaneous functions