---
title: "TidyData"
author: "Ali Paulson"
date: "February 13, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
#Can use warning = FALSE, message = FALSE to suppress warnings in the html output
```


The data we are using are from [Mike Byerly. Alaska commercial salmon catches by management region (1886- 1997). Gulf of Alaska Data Portal. df35b.304.2.](https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1) This is from a PDF file, retrieved with OCR - optical character recognition. You can use this to get a computer to read a PDF. 

```{r data, echo = FALSE}
#catch <- read.csv("../salmondata/df35b.302.1.csv", stringsAsFactors = FALSE)

#regdefs <- read.csv("../salmondata/df35b.303.1.csv", stringsAsFactors = FALSE)

#Can also just download these from the internet!
catch_original <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method = "libcurl"), stringsAsFactors = FALSE)
```

# Reformat Catch Data

```{r}
head(catch_original)
```

Remove the `all` and `notesRegCode` columns

Some people just remove columns by setting them to NULL, but this isn't very clean/readable.  The dplyr world makes things more readable, etc. So, we can use **select()** to remove columns. 

The pipe operator takes the output from one function and puts it into the input of the next function. <- is kind of like "gets the value", %>% is like "gets passed on to/then". Piping makes your workflow more legible, easier for others to understand. 

Ctrl + shift + m: pipe operature shortcut %>% 

Need to explain why you don't have to write in the dataframe as the first argument as select...pipe operator passes things into that first argument. 

**gather()** collects things into key-value pairs

* key is the column names in wide format - key specifies what we want to call the new column of old column names
* value is the value in those columns - value is what we want to call those values in the new long dataframe
* gather() will turn all of your columns into these key-value pairs unless you tell it otherwise
    + can tell gather() what not to include (or what to include)
        + e.g. -Region, -Year
        + or e.g. Chinook, Sockeye, Coho, Pink, Chum

```{r}

catch_long <- catch_original %>% 
  select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum)

head(catch_long)

#equivalent to
# catch_longb <- catch_original %>% 
#   select(-All, -notesRegCode)
# head(catch_longb)

#Now let's make it long
catch_long <- catch_original %>% 
  #Select desired columns
  select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum) %>% 
  #Move from wide to long
  gather(key = "species", value = "catch", -Region, -Year)

head(catch_long)
```


The opposite of **gather()** is **spread()**. 

* key - the column names - this is the column that you want to turn into column names  
* value is the column with the values that you want the new df to have

```{r}
catch_wide <- catch_long %>% 
  spread(key = species, value = catch)
head(catch_wide)

```


* Create species column
    + move from wide to long
* General QA

# Join the Region Definitions 

# Miscellaneous functions